### 2014-02-25 phpmyadmin/GSoC_infosession_05Mar2014/presentation.tex
update date and stats

\documentclass[14pt]{beamer}
\usepackage{hyperref}
\usepackage[T1]{fontenc}
\usetheme{Madrid}
\definecolor{PmaPurple}{RGB}{00,33,66}
\definecolor{PmaOrange}{RGB}{255,173,17}
\definecolor{PmaOlive}{HTML}{727E0A}
\usecolortheme[named=PmaPurple]{structure}
\setbeamercolor{title}{fg=PmaOrange}
\setbeamercolor{frametitle}{fg=PmaOrange}
\hypersetup{colorlinks,linkcolor=,urlcolor=PmaOrange}
\logo{
  \includegraphics[scale=.6]{pictures/pma_logo.png}
}
\title{GSoC 2014 info session - phpMyAdmin}
\author{Dieter Adriaenssens}
\institute[phpMyAdmin]{phpMyAdmin GSoC mentor 2011-2013}
\date[5Mar2014]{Google Summer of Code info session UGent\\
Mentoring organisation : phpMyAdmin\\
March 5th, 2014}
\subject{phpMyAdmin}
\begin{document}
  \begin{frame}
  \titlepage
  \end{frame}
  \begin{frame}
    \frametitle{What is phpMyAdmin?}
    %(question to audience : who knows/uses pma?)
      A webinterface written in PHP (and JavaScript),\\
      to administer a MySQL database and users.
  \end{frame}
  \begin{frame}
    \frametitle{Current version}
    % (questions to audience about used version)
    \begin{itemize}
      \item stable version : 4.1.8
      \item in development : 4.2
      \item available in \href{http://www.phpmyadmin.net/home\_page/translations.php}{72 languages}
      \item avg. 100K downloads (every release)
      \item 3M downloads in total last year
      \item 10K commits by 285 contributors last year
      \item approx. 20 active contributors every month
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{Why phpMyAdmin participates in GSoC}
    \begin{itemize}
      \item to get new features implemented in future versions
      \item attract new developers
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{GSoC history}
    \begin{itemize}
      \item 2012 : 6 students, 5 passed -> phpMyAdmin 4.0
      \item 2011 : 8 students, 7 passed -> phpMyAdmin 3.5
      \item 2010 : 6 students, 6 passed -> phpMyAdmin 3.4
      \item 2009 : under MySQL umbrella : 5 students
      \item 2008 : under MySQL umbrella : 2 students
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{What made it into 3.5}
    \begin{itemize}
      \item improved support for stored routines, events and triggers
      \item openGIS support
      \item Drizzle support
      \item improved ENUM/SET editor
      \item advisor : suggestions to tweak config setting in order to optimise MySQL server performance
      \item zoom search
      \item grid editing
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{What made it into 4.0}
    \begin{itemize}
      \item code cleanup and refactoring
      \item plugin system for import/export/auth/\ldots
      \item removal of frames
      \item new navigation tree
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{Ideas list for GSoC 2013}
    Full list : \href{http://wiki.phpma.org/pma/GSoC\_2013\_Ideas\_List}{wiki.phpma.org/pma/GSoC\_2013\_Ideas\_List}
    \begin{itemize}
      \item automated testing
      \item refactoring
      \item synchronize data
      \item AJAX error reporting
      \item interface improvements
      \item your idea\ldots
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{What to expect}
    \begin{itemize}
      \item mentoring by experienced mentors/developers
      \item be active on the mailinglist
      \item submit a patch fixing a bug or adding a small feature
      \item 12 weeks of 40 hours coding
    \end{itemize}
 \end{frame}
  \begin{frame}
    \frametitle{Submitting a good proposal}
    \begin{itemize}
      \item write a good project proposal :
      \begin{itemize}
        \item detailed functional description
        \item well defined deliverables
        \item clear schedule
      \end{itemize}
      \item submit well in advance : time for feedback
      \item application deadline : May 3
      \item patch deadline (phpMyAdmin) : May 10
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{During GSoC}
    \textbf{May 27} - When you get selected :
    \begin{itemize}
      \item community bonding period : get up to speed, read, ask questions, set up your wiki, clone the git repo, Google administration
      \pause
      \item \textbf{June 17} - start coding, first payment
      \item mandatory weekly reports on your blog, commit code regularly, merged into master by mentor
      \pause
      \item \textbf{August 2} - mid-term evaluation, based on your schedule + payment
      \pause
      \item finish coding, documentation, testing
      \pause
      \item \textbf{September 16} - pencils down, final evaluation, submit code, final payment
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{What happens after GSoC}
    \begin{itemize}
      \item your code gets merged with master branch
      \item we offer active membership :
      \begin{itemize}
        \item commit access
        \item team membership
        \item invitation to team meetings
      \end{itemize}
    \end{itemize}
  \end{frame}
  \begin{frame}
   \frametitle{Questions}
    Thanks for your attention!\\
    Questions?
    \begin{itemize}
      \item website : \href{http://www.phpmyadmin.net/}{www.phpmyadmin.net}
      \item GSoC info : \href{http://wiki.phpma.org/pma/GSoC\_2013\_Applicant\_Guide}{wiki.phpma.org/pma/GSoC\_2013\_Applicant\_Guide}
      \item developers mailing list (development discussion): \href{mailto:phpmyadmin-devel@lists.sourceforge.net}{phpmyadmin-devel@lists.sourceforge.net}
    \end{itemize}
  \end{frame}
\end{document}