### 2014-02-26 phpmyadmin/GSoC_infosession_05Mar2014/presentation.tex
update links, dates and ideas to 2014

\documentclass[14pt]{beamer}
\usepackage{hyperref}
\usepackage[T1]{fontenc}
\usetheme{Madrid}
\definecolor{PmaPurple}{RGB}{00,33,66}
\definecolor{PmaOrange}{RGB}{255,173,17}
\definecolor{PmaOlive}{HTML}{727E0A}
\usecolortheme[named=PmaPurple]{structure}
\setbeamercolor{title}{fg=PmaOrange}
\setbeamercolor{frametitle}{fg=PmaOrange}
\hypersetup{colorlinks,linkcolor=,urlcolor=PmaOrange}
\logo{
  \includegraphics[scale=.6]{pictures/pma_logo.png}
}
\title{GSoC 2014 info session - phpMyAdmin}
\author{Dieter Adriaenssens}
\institute[phpMyAdmin]{phpMyAdmin GSoC mentor 2011-2013}
\date[5Mar2014]{Google Summer of Code info session UGent\\
Mentoring organisation : phpMyAdmin\\
March 5th, 2014}
\subject{phpMyAdmin}
\begin{document}
  \begin{frame}
  \titlepage
  \end{frame}
  \begin{frame}
    \frametitle{What is phpMyAdmin?}
    %(question to audience : who knows/uses pma?)
      A webinterface written in PHP (and JavaScript),\\
      to administer a MySQL database and users.
  \end{frame}
  \begin{frame}
    \frametitle{Current version}
    % (questions to audience about used version)
    \begin{itemize}
      \item stable version : 4.1.8
      \item in development : 4.2
      \item available in \href{http://www.phpmyadmin.net/home\_page/translations.php}{72 languages}
      \item avg. 100K downloads (every release)
      \item 3M downloads in total last year
      \item 10K commits by 285 contributors last year
      \item approx. 20 active contributors every month
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{Why phpMyAdmin participates in GSoC}
    \begin{itemize}
      \item get new features implemented in future versions
      \item attract new developers
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{GSoC history}
    \begin{itemize}
      \item 2013 : 6 students, 6 passed -> phpMyAdmin 4.1
      \item 2012 : 6 students, 5 passed -> phpMyAdmin 4.0
      \item 2011 : 8 students, 7 passed -> phpMyAdmin 3.5
      \item 2010 : 6 students, 6 passed -> phpMyAdmin 3.4
      \item 2009 : under MySQL umbrella : 5 students
      \item 2008 : under MySQL umbrella : 2 students
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{What made it into 3.5}
    \begin{itemize}
      \item improved support for stored routines, events and triggers
      \item openGIS support
      \item Drizzle support
      \item improved ENUM/SET editor
      \item advisor : suggestions to tweak config setting in order to optimise MySQL server performance
      \item zoom search
      \item grid editing
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{What made it into 4.0}
    \begin{itemize}
      \item code cleanup and refactoring
      \item plugin system for import/export/auth/\ldots
      \item removal of frames
      \item new navigation tree
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{Ideas list for GSoC 2014}
    Full list : \href{http://wiki.phpma.org/pma/GSoC\_2014\_Ideas\_List}{wiki.phpma.org/pma/GSoC\_2014\_Ideas\_List}
    \begin{itemize}
      \item automated testing
      \item (user) interface improvements
      \item Error reporting
      \item structure tools 
      \begin{itemize}
        \item database normalization
        \item central list of columns
      \end{itemize}
      \item your idea\ldots
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{What to expect}
    \begin{itemize}
      \item mentoring by experienced mentors/developers
      \item be active on the mailinglist
      \item submit a patch fixing a bug or adding a small feature
      \item 12 weeks of 40 hours coding
    \end{itemize}
 \end{frame}
  \begin{frame}
    \frametitle{Submitting a good proposal}
    \begin{itemize}
      \item write a good project proposal :
      \begin{itemize}
        \item detailed functional description
        \item well defined deliverables
        \item clear schedule
      \end{itemize}
      \item submit well in advance : time for feedback
      \item \textbf{March 21st} : application deadline
      \item \textbf{March 28th} : patch deadline (phpMyAdmin)
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{During GSoC}
    When you get selected :
    \begin{itemize}
      \item community bonding period : get up to speed, read, ask questions, set up your wiki, clone the git repo, Google 
      \item mandatory weekly reports on your blog, commit code regularly, merged into master by mentor
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{What happens after GSoC}
    \begin{itemize}
      \item your code gets merged with master branch
      \item we offer active membership :
      \begin{itemize}
        \item commit access
        \item team membership
        \item invitation to team meetings
      \end{itemize}
    \end{itemize}
  \end{frame}
  \begin{frame}
   \frametitle{Questions}
    Thanks for your attention!\\
    Questions?
    \begin{itemize}
      \item website : \href{http://www.phpmyadmin.net/}{www.phpmyadmin.net}
      \item GSoC info : \href{http://wiki.phpma.org/pma/GSoC\_2014\_Applicant\_Guide}{wiki.phpma.org/pma/GSoC\_2014\_Applicant\_Guide}
      \item developers mailing list (development discussion): \href{mailto:phpmyadmin-devel@lists.sourceforge.net}{phpmyadmin-devel@lists.sourceforge.net}
    \end{itemize}
  \end{frame}
\end{document}