### 2014-03-24 iptables_newline2014/presentation.tex
create presentation about iptables

\documentclass[14pt]{beamer}
\usepackage{hyperref}
\usepackage[T1]{fontenc}
\usetheme{Madrid}
%\usecolortheme[named=PmaPurple]{structure}
%\setbeamercolor{title}{fg=PmaOrange}
%\setbeamercolor{frametitle}{fg=PmaOrange}
%\hypersetup{colorlinks,linkcolor=,urlcolor=PmaOrange}
\title[Reduce iptables config complexity]{Reducing iptables configuration complexity using chains}
\author{Dieter Adriaenssens}
%\institute[phpMyAdmin]{phpMyAdmin developer}
\date[29Mar2014]{Newline 0x04\\
Whitespace, Ghent\\
March 29th, 2014}
\subject{iptables}
\begin{document}
  \begin{frame}
  \titlepage
  \end{frame}
  \begin{frame}
    \frametitle{What is iptables}
    %(question to audience : who knows/uses iptables?)
  \end{frame}
  \begin{frame}
   \frametitle{Questions}
    Thanks for your attention!\\
    Questions?\\
    Contact
    \begin{itemize}
      \item Blog : \href{http://ruleant.blogspot.com/}{http://ruleant.blogspot.com}
      \item Twitter : \href{https://twitter.com/dcadriaenssens}{@dcadriaenssens}
    \end{itemize}
  \end{frame}
\end{document}