### 2015-03-23 buildtimetrend/presentation.tex
update title and links

\documentclass[14pt]{beamer}
\usepackage{hyperref}
\usepackage{ulem}
\usepackage[T1]{fontenc}
\usetheme{Madrid}

% macro for double dash
\newcommand{\dd}{{\texttt{-{}-}}}

%% Enable license helpers
\graphicspath{{../cc_beamer/}}
\input{../cc_beamer/cc_beamer}

\title[Buildtime Trend : What, why and how]{Buildtime Trend : What's trending in your build process}
\author{Dieter Adriaenssens}
\institute[Buildtime Trend]{@dcadriaenssens - Founder, developer Buildtime Trend}
\date[NewLine 5Apr2015]{NewLine - Ghent\\
April 5th, 2015}
\subject{Buildtime Trend}
\begin{document}
  \begin{frame}
    \titlepage
    \vfill
    \begin{center}
      \CcGroupByNcSa{0.83}{0.95ex}\\[2.5ex]
        {\tiny\CcNote{\CcLongnameByNcSa}}
        \vspace*{-2.5ex}
    \end{center}
  \end{frame}
  \begin{frame}
    \frametitle{Overview}
    \begin{itemize}
      \item What is Buildtime Trend
      \item Why and how
      \item Demo
      \item Lessons learned
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{What is Buildtime Trend?}
    %(question to audience : who knows/uses iptables?)
  \end{frame}
  \begin{frame}
    \frametitle{Lessons learned}
    \begin{itemize}
      \item keep it simple
    \end{itemize}
  \end{frame}
  \begin{frame}
   \frametitle{Questions}
    Thanks for your attention!\\
    Questions?
    \vfill
    Contact
    \begin{itemize}
      \item Website : \href{https://buildtimetrend.github.io/}{https://buildtimetrend.github.io/}
      \item Twitter : \href{https://twitter.com/buildtime_trend}{@buildtime\_trend}
    \end{itemize}
  \end{frame}
\end{document}