### 2015-04-04 oss365/presentation.tex
start oss365 presentation

\documentclass[14pt]{beamer}
\usepackage{hyperref}
\usepackage{ulem}
\usepackage[T1]{fontenc}
\usetheme{Madrid}

%% Enable license helpers
\graphicspath{ {../cc_beamer/}{pictures/} }
\input{../cc_beamer/cc_beamer}

\title[365 days of Open Source]{What I learned from 365 days of contributing to Open Source software}
\author{Dieter Adriaenssens}
\institute[]{@dcadriaenssens}
\date[NewLine 0x05 5Apr2015]{NewLine 0x05 - Ghent\\
April 5th, 2015}
\subject{365 days of Open Source}
\begin{document}
  \begin{frame}
    \titlepage
    \vfill
    \begin{center}
      \CcGroupByNcSa{0.83}{0.95ex}\\[2.5ex]
        {\tiny\CcNote{\CcLongnameByNcSa}}
        \vspace*{-2.5ex}
    \end{center}
  \end{frame}
  \begin{frame}
    \frametitle{365 day of Open Source}
    \begin{itemize}
      \item Github : 365 day commit streak, 2332 commits
      \item \href{https://github.com/ruleant}{Today : 447 day commit streak} (and counting)
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{How it started}
    \begin{itemize}
      \item 14 January 2014 : returned from 'unplugged' holiday
      \item previous 100+ day commit streaks
      \item see how far I would get
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{June 2014 : Getting serious}
    \begin{itemize}
      \item Josh (@dzello) of Keen.io pledged to do a 365 day commit streak
      \item I was at 162 days 
      \item I took the challenge to do a 365 day streak
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{How}
    \begin{itemize}
      \item working on my projects : Buildtime Trend, GetBack GPS
      \item new features 
      \item refactoring
      \item add unit tests
      \item coding style
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{Lessons learned}
    \begin{itemize}
      \item void
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{Acknowledgements}
    Big thanks to
    \begin{itemize}
      \item Josh (\href{https://twitter.com/dzello}{@dzello}) of Keen.io
      \item my friends and family
    \end{itemize}
  \end{frame}
  \begin{frame}
   \frametitle{Questions}
    Thanks for your attention!\\
    Questions?
    \vfill
    Dieter Adriaenssens - \href{https://twitter.com/dcadriaenssens}{\small{@dcadriaenssens}}
  \end{frame}
\end{document}