### 2015-04-04 oss365/presentation.tex
add more slides

\documentclass[14pt]{beamer}
\usepackage{hyperref}
\usepackage{ulem}
\usepackage[T1]{fontenc}
\usetheme{Madrid}

%% Enable license helpers
\graphicspath{ {../cc_beamer/}{pictures/} }
\input{../cc_beamer/cc_beamer}

\title[365 days of Open Source]{What I learned from 365 days of contributing to Open Source software}
\author{Dieter Adriaenssens}
\institute[]{@dcadriaenssens}
\date[NewLine 0x05 5Apr2015]{NewLine 0x05 - Ghent\\
April 5th, 2015}
\subject{365 days of Open Source}
\begin{document}
  \begin{frame}
    \titlepage
    \vfill
    \begin{center}
      \CcGroupByNcSa{0.83}{0.95ex}\\[2.5ex]
        {\tiny\CcNote{\CcLongnameByNcSa}}
        \vspace*{-2.5ex}
    \end{center}
  \end{frame}
  \begin{frame}
    \frametitle{365 day of Open Source}
    \begin{itemize}
      \item Github : 365 day commit streak, 2332 commits
      \item \href{https://github.com/ruleant}{Today : 447 day commit streak} (and counting)
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{How it started}
    \begin{itemize}
      \item 14 January 2014 : returned from 'unplugged' holiday
      \item previous 100+ day commit streaks
      \item see how far I would get
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{June 2014 : Getting serious}
    \begin{itemize}
      \item Josh (@dzello) of Keen.io pledged to do a 365 day commit streak
      \item I was at 162 days 
      \item I took the challenge to do a 365 day streak
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{How}
    \begin{itemize}
      \item working on my projects : Buildtime Trend, GetBack GPS
      \item new features 
      \item refactoring
      \item add unit tests
      \item coding style
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{Holiday}
    \begin{itemize}
      \item July 2014 : climbing trip
      \item climbing during the day
      \item limited time in the evening
      \item planning to do small commits
      \item close call on a few days
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{Difficulties}
    \begin{itemize}
      \item Traveling
      \item Tired after a long day at work
      \item stuck on a problem
    \end{itemize}
    Will I keep my streak going?
  \end{frame}
  \begin{frame}
    \frametitle{Meeting Josh}
    \begin{itemize}
      \item Meeting Josh in San Francisco in October 2014
      \item Josh stopped his 66 day streak going to Burning Man
      \item other ways of measuring progress
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{Lessons learned}
    \begin{itemize}
      \item set a realistic goal
      \item choose something that works for you
      \item minimum 1 commit per day worked for me
      \item I will keep the streak going as long as I am comfortable with
      \item working on the project should be the primary goal
      \item pledge to commit every day is means to keep you going
      \item or any sort of metrics
      \item similarities in sports, goals, projects, \ldots
      \item allow to give yourself a break
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{Acknowledgements}
    Big thanks to
    \begin{itemize}
      \item Josh (\href{https://twitter.com/dzello}{@dzello}) of Keen.io
      \item my friends and family
    \end{itemize}
  \end{frame}
  \begin{frame}
   \frametitle{Questions}
    Thanks for your attention!\\
    Questions?
    \vfill
    Dieter Adriaenssens - \href{https://twitter.com/dcadriaenssens}{\small{@dcadriaenssens}}
    \vfill
    Blogpost : \href{http://ruleant.blogspot.be/2015/01/what-i-learned-from-365-days-of.html}{http://ruleant.blogspot.be/2015/01/what-i-learned-from-365-days-of.html}
  \end{frame}
\end{document}